### 2025-09-09 age_structure_forecasts.qmd
Added more discussion

---
title: "Forecasting the age structure of the scientific workforce in Australia"
author:
- name: Rob J Hyndman
  affiliations:
    - name: Monash University
      department: Department of Econometrics & Business Statistics
      city: Clayton
      country: Australia
      postal-code: 3800
  email: Rob.Hyndman@monash.edu
  corresponding: true
- name: Kelly Vanh Nguyen
  affiliations:
    - name: Monash University
      department: Department of Econometrics & Business Statistics
      city: Clayton VIC
      country: Australia
      postal-code: 3800
abstract: |
  Planning for a future workforce requires forecasts of age structure changes to inform policy decisions, particularly related to universities and immigration. We modify the traditional demographic growth-balance equation to account for workforce-specific dynamics, replacing births with graduate entry, modeling exits through death and retirement, and including a remainder term that captures migration and career changes. Functional data models are used to model age-specific components, while ARIMA models are used for time series components. Simulation is employed to generate forecast distributions, capturing uncertainty from all components. The approach is illustrated using data on Australia's scientific workforce, allowing us to forecast the age distribution of various scientific disciplines for the next ten years. This analysis was central to an Australian Academy of Science initiative, supporting strategic workforce planning and policy development for Australia's scientific community.
keywords:
  - cohort analysis
  - demographic modelling
  - functional data models
  - labour market
  - workforce planning
bibliography: refs.bib
wpnumber: no/yr
jelcodes: J11, J21, C15, C22
blind: false
cover: true
linestretch: 1.5
include-in-header: preamble.tex
format:
  wp-pdf:
    keep-tex: true
filters:
  - blindable-div
---

```{r setup, include=FALSE}
library(targets)
library(ggplot2)

# Set default colors
options(
  ggplot2.discrete.colour = c(
    "#D55E00",
    "#0072B2",
    "#009E73",
    "#CC79A7",
    "#E69F00",
    "#56B4E9",
    "#F0E442"
  ),
  ggplot2.discrete.fill = c(
    "#D55E00",
    "#0072B2",
    "#009E73",
    "#CC79A7",
    "#E69F00",
    "#56B4E9",
    "#F0E442"
  )
)
```

# Introduction {#sec-intro}

In planning for the future labour market, it is necessary to forecast the age structure of the workforce in order to enable informed decision-making on policies, especially concerning universities and immigration. We propose a statistical modelling approach to this problem, illustrated using various scientific disciplines in Australia, forecasting future workforce age structures over the next decade. The forecasts described have been used by the Australian Academy of Science as part of _Australian Science, Australia's Future: Science 2035_, an initiative assessing the capability of the national science system and its role in achieving Australia's ambitions [@AAS2035].

The economic implications of workforce age structure shifts are well-documented [e.g., @Bloom2007], affecting productivity, pensions and superannuation, and skill shortages [@PC2013; @OECD2019a; @OECD2019b; @pension]. The social implications are also significant, with an aging workforce leading to changes in workplace dynamics, potential problems with intergenerational knowledge transfer, and the need for policies that support older workers. Yet this problem does not appear to have been previously addressed from a statistical modelling perspective.

Our approach builds on functional data models, introduced to demographic modelling by @HU07. They combined nonparametric smoothing and functional principal components for age-specific demographic rates. These models were then used by @HB08 for mortality, fertility, and migration rates, providing stochastic data generating processes for the components of demographic balance equations. These separate component models were then simulated to form future sample paths, leading to age- and sex-specific stochastic population forecasts. The modelling framework was later extended by @HBY13 to ensure coherence of forecasts between sexes or other demographic groups.

We propose to adapt this framework to workforce dynamics by redefining the demographic components in two ways. First, we replace fertility with workforce entry, which functions more like a migration process than a birth process because graduates can enter the workforce at any age. Second, we allow workers to leave the workforce through two processes: retirement and death. Of course, people may also leave the workforce for other reasons, such as a career change or family commitments, but since we do not have data on these processes, we model them implicitly via a remainder term.

We describe the methodology in @sec-methodology. By way of illustration, we apply the methodology to major scientific disciplines in Australia, focusing on the Natural and Physical Sciences. We describe the data sources in @sec-data, with the results provided in @sec-results. The aim of this analysis is to inform future workforce planning and policy decisions to support the growth of Australia's scientific community. Finally, we provide some discussion and conclusions in @sec-conclusion.

# Methodology {#sec-methodology}

Suppose our workforce is divided into $I$ groups, indexed by $i=1,\dots,I$. In our application, these are scientific disciplines, but in principle they could refer to any subdivision of workers. Let $P_{i,x,t}$ denote the number of equivalent full-time workers in group $i$ who are aged $x$ at the start of year $t$, where $x=15,16,\dots$. The starting age of 15 is chosen because it is the minimum age at which individuals are counted as part of the labour force in the Australian Census [@ABSLFSP]. We assume that data are available for years $t=1,\dots,T$, and that forecasts are required for $P_{i,x,T+h}$ across all ages and groups, for some forecast horizon $h>0$.

People can leave the workforce of a group through death, retirement, emigration, or career change; they can enter the workforce through graduation, immigration, or career change. Unfortunately, we typically do not have data on many of these processes, so we will combine career change, emigration and immigration into a remainder term, which we denote as $E_{i,x,t}$. Let $D_{i,x,t}$ denote the number of deaths of workers in group $i$ of age $x$ in year $t$, $R_{i,x,t}$ denote the number of retirements from the same group of workers, and $G_{i,x,t}$ denote the number of new graduates of age $x$ in year $t$ who take up work in group $i$. The numbers in each case are for people aged $x$ at the *start* of year $t$. Then population changes can be described using a model similar to the stochastic population model of @HB08, excluding the birth process:
$$
P_{i,x+1,t+1} = P_{i,x,t} - D_{i,x,t} - R_{i,x,t} + G_{i,x,t} + E_{i,x,t},
$$ {#eq-dgb}
where

* $D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{i,x,t})$, with $q_{i,x,t}$ being the probability of death for group $i$ at age $x$ in year $t$; and
* $R_{i,x,t} \sim \text{Binomial}((P_{i,x,t}-D_{i,x,t}), r_{i,x,t})$, with $r_{i,x,t}$ being the probability of retirement from group $i$ at age $x$ in year $t$.

That is, the population each year is equal to the population from the previous year having aged 1 year, minus the deaths or retirements that occurred during the previous year, plus the new graduates, plus any other changes due to migration or career change (which may be negative). We assume that $E_{i,x,t} = G_{i,x,t} = 0$ above some age threshold (say $x=100$). Once $P_{i,x,t} = 0$ when $x$ is above that threshold, all future populations $P_{i,x+k,t+k} = 0$, for $k=1,2,\dots$. That is, when the cohort aged $x$ in year $t$ has all retired or died, and $x$ is above the threshold, they will not be replaced by new workers of the same age.

As a first approximation, the components $q$, $r$, $E$ and $G$ can be assumed to behave independently for each combination of $i$, $x$ and $t$. In reality, there may be some negative correlation between $G$ and $E$ as insufficient graduates would probably lead to employers finding people from overseas, while too many graduates would lead to scientists seeking work elsewhere.

The choice of a Binomial rather than a Poisson distribution (in contrast to @Brillinger1986) for deaths and retirements is because the Binomial distribution ensures that the number of deaths and retirements cannot exceed the population at risk. In a simulation context, with very small populations, this is important to avoid nonsensical results.

It is unlikely that we have available separate death and retirement counts for each group, and retirement data is not available in all years. So we will let $q_{i,x,t} = q_{x,t}$ and $r_{i,x,t} = r_{x}$, assuming that death rates and retirement rates are the same across all groups, and that retirement rates do not change over time. Similarly, graduation numbers are rarely available by discipline and age, so we will approximate $G_{i,x,t} = g_xG_{i,t}$ where $G_{i,t}$ is the total number of graduates in year $t$ and $g_x$ is the proportion of graduates by age across all disciplines.

This leads to the simpler model
$$
P_{i,x+1,t+1} = P_{i,x,t} - D_{i,x,t} - R_{i,x,t} + g_xG_{i,t} + E_{i,x,t},
$$ {#eq-dgb2}
where

* $D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})$; and
* $R_{i,x,t} \sim \text{Binomial}((P_{i,x,t}-D_{i,x,t}), r_{x})$.

For the age-specific time-varying components $q_{x,t}$ and $E_{i,x,t}$, we will use functional data models [@HU07]. For $q_{x,t}$, this is of the form
$$
\log(q_{x,t}) = \mu(x) + \sum_{k=1}^K \beta_{k,t} \phi_k(x) + \varepsilon_{x,t},
$$ {#eq-fdm}
where $\mu(x)$ is the mean function, $\phi_k(x)$ are functional principal components, $\beta_{k,t}$ are the principal component scores, and $\varepsilon_{x,t}$ is an error term. Each $\beta_{k,t}$ is then modelled using a univariate time series model, such as an ARIMA model. A log-link function is used to ensure that the probabilities remain positive. An inverse logit link function could also be used, if the probabilities are close to 1 for some ages. A similar model is used for $E_{i,x,t}$, with separate mean functions and principal components for each group $i$. The number of components $K=6$, for the reasons outlined in @HB08.

Functional data models have been widely used in demography and other fields, and have been shown to work particularly well for age-specific demographic processes [@HB08;@LCcomparison]. They enable the inherent smoothness over age to be captured, while modelling the autocorrelation over time using relatively simple univariate time series models applied to the principal component scores. In our application, we use univariate ARIMA models for the $q_{x,t}$ scores, and ARMA models for the $E_{i,x,t}$ scores. The assumption of stationarity for the $E_{i,x,t}$ scores is validated for the disciplines we consider.

For the time-varying component, $G_{i,t}$, we use a global ARIMA model [@HM21] to capture the dynamics over time and across disciplines. The global model pools information across disciplines to improve forecast accuracy, especially for disciplines with limited historical data.

To forecast future working population numbers, $P_{i,x,t}$, $t>T$, we simulate future sample paths of each of the components $G_{i,t}$, $q_{x,t}$, and $E_{i,x,t}$, simulate $D_{i,x,t}$ and $R_{i,x,t}$ from their respective Binomial distributions, and then use the demographic growth-balance equation @eq-dgb2 iteratively to obtain $P_{i,x,t}$ for $t=T+1,T+2,\dots$. This simulation-based approach allows us to capture the uncertainty in each of the components, leading to a distribution of possible future outcomes for $P_{i,x,t}$.

This model is somewhat pragmatic given the data available in our specific application. If better data were available, other variations on @eq-dgb could be used. For example, if death rates were available by discipline, then we would replace $q_{x,t}$ by $q_{i,x,t}$ in the Binomial deaths distribution. If retirement rates were available by year, or by discipline, we could similarly replace $r_x$ by a more specific retirement rate in the Binomial retirements distribution. If we had data on graduations by age and discipline, we could replace $g_xG_{i,t}$ by $G_{i,x,t}$. If we had data on migration, we could split the remainder $E_{i,x,t}$ into several components, and model them separately. None of this changes the overall modelling framework we are proposing.

# Data {#sec-data}

To illustrate the methodology, we consider the Natural and Physical Sciences as defined in the Australian Standard Classification of Education (ASCED) by the @ABSASCED. We refer to ASCED's Narrow Fields as "disciplines"; these comprise Physics and Astronomy, Mathematical Sciences, Chemical Sciences, Earth Sciences, Biological Sciences, Other Natural and Physical Sciences, and Natural and Physical Sciences not further defined (n.f.d.). @tbl-disciplines lists the detailed fields within each scientific discipline.

```{r}
#| label: tbl-disciplines
#| tbl-cap: Classification of scientific disciplines, based on the ASCED Narrow Fields of Education within the Broad Field of Natural and Physical Sciences. The table lists their corresponding Detailed Fields.  "n.e.c." stands for "Not Elsewhere Classified."
tar_read(discipline_table)
```

We define the population of workers in a discipline as those who are active in the labour market and hold a bachelor's degree or higher in that discipline. For the purposes of this analysis, we will omit "Other Natural and Physical Sciences" and "Natural and Physical Sciences n.f.d.".

## Working population

Data on the working population were sourced from the *Census of Population and Housing* [@ABSCensus] for census years 2006, 2011, 2016, and 2021. This dataset encompasses one-year age groups, the highest level of completed non-school qualification level (QALLP), the corresponding field of study [QALFP, @ABSQALFP], and the industries in which individuals work. However, labour force participation status [@ABSLFFP] is available only for 2016 and 2021. To estimate worker numbers for 2006 and 2011, the average participation rates from 2016 and 2021 were applied, assuming overall age distributions remain consistent.

The resulting estimates of the number of scientists who are active in the Australian labour market is shown in @fig-cohort-interpolation as the thick lines. Cohort interpolation [@STUPP88], applying linear interpolation within each age cohort between census years, is used to estimate values for the intercensal years (shown as thin lines), giving $P_{i,x,t}$ for each discipline $i$, age $x$, and year $t$.

```{r}
#| label: fig-cohort-interpolation
#| fig-cap: "$P_{i,x,t}$: Estimated number of working scientists in Australia by discipline and age, 2006--2021. Thicker lines are used to denote census years."
tar_read(fig20)
```

## Retirements

Retirement data was sourced from the *Retirement and Retirement Intentions* dataset (Catalogue 6238) for the 2022--2023 financial year [@ABS6238]. The data are categorised by the industry of an individual's main job, and are provided in four broad age groups (45--59, 60--64, 65--69 and 70+). There are 19 industry categories, with the largest numbers of scientists working in Education and Training (15.8%), Professional, Scientific and Technical Services (15.5%), and Health Care and Social Assistance (14.6%). The proportions in other industries are much smaller. We take a weighted average of retirement intentions using these top three industries, with proportions rescaled to sum to 1. The resulting values are shown in @fig-retirements as the gray line. To obtain a single-year-of-age retirement distribution, we disaggregate the data using a monotonic cubic spline applied to the cumulative values of these age groups [@SHW04]. The resulting smoothed distribution ($r_x$) is shown as the black line in @fig-retirements.

```{r}
#| label: fig-retirements
#| fig-cap: "$r_x$: Age distribution of retirement intentions, based on data from the 2022--2023 Australian financial year. The blue line shows the smoothed probabilities."
tar_read(fig_rx2)
```

## Deaths

Age-specific mortality rates from 1971 to 2021 were obtained from the @HMD. Using standard life table methods, these rates are converted into age-specific probabilities of death, as shown in @fig-death-probs. Over time, mortality probabilities have generally declined across all age groups, reflecting improvements in Australian life expectancy.

```{r}
#| label: fig-death-probs
#| fig-cap: "$q_{x,t}$: Age-specific probabilities of death (on a logarithmic scale) for each year from 1971 to 2021."
tar_read(fig_mxt)
```

No data are available for specific industry groups, so we assume that all scientists have the same mortality probabilities as the general population. These probabilities serve as estimates of $q_{x,t}$.

## Graduate completions

Graduate completion statistics were obtained from the *Award Course Completions* dataset [@DeptEdu]. @fig-completions shows the distribution of graduate completions with a bachelor's degree or higher, by age for each year from 2006 to 2023. Some missing values result in gaps in certain lines, but the overall pattern remains highly consistent across years. Given this consistency, the data is averaged across all available years, and then smoothed by applying monotonic cubic splines to the cumulative values [@SHW04]. The resulting averaged distribution, shown as the black line in @fig-completions, smooths out year-to-year fluctuations and provides an  estimate of $g_x$.

```{r}
#| label: fig-completions
#| fig-cap: "$g_x$: Estimated distribution of graduate completions by age (black). This is estimated by averaging and smoothing the data for the years 2006 to 2023 (coloured)."
tar_read(fig_ave_completions)
```

The Department of Education provides data on the number of graduates with a bachelor's degree or higher, categorised by discipline and year [@DeptEdu_Private]. This dataset includes both domestic and international students. The total number of graduates, $G_{i,t}$, in each discipline $i$ and year $t$, are shown in @fig-graduates.

```{r}
#| label: fig-graduates
#| fig-cap: "$G_{i,t}$: Total number of graduates with a bachelor's degree or higher by discipline from 2006 to 2023."
tar_read(fig21)
```

The large increase in the working population observed in the 2021 Census for Mathematical Sciences (@fig-cohort-interpolation) can be partly attributed to the sharp rise in graduate numbers between 2016 and 2021. This is probably due to the impact of data science, and the growing importance of statistics and machine learning in many areas of employment.

## Remainder

The demographic growth-balance equation (@eq-dgb2), when rearranged, provides an expression for the remainder including net migration and career changes:
$$
E_{i,x,t} = P_{i,x+1,t+1} - P_{i,x,t} - D_{i,x,t} - R_{i,x,t} - g_xG_{i,t},
$$ {#eq-remainder}
However, we do not have data on $D_{i,x,t}$ and $R_{i,x,t}$, so we replace these by their expected values, $P_{i,x,t} q_{x,t}$ and $P_{i,x,t}(1-q_{x,t}) r_x$, respectively. We can only estimate remainders up to 2020 because we need data for both year $t$ and year $t+1$ in @eq-remainder, and our working population data only extends to 2021. The estimated remainders are shown in @fig-remainder.

```{r}
#| label: fig-remainder
#| fig-cap: Estimated remainder $E_{i,x,t}$ by discipline, age and year (2006--2020).
tar_read(fig22)
```

The inclusion of international students in the graduate data leads to large positive values of the remainder for the teenage years, followed by large negative values when these students return to their home countries after graduation.

# Results {#sec-results}

## Graduate completions

To forecast future graduate numbers, $G_{i,t}$, a global ARIMA model was employed, following the principles outlined by @HM21. The global model captures overall trends across disciplines by scaling graduate data within each discipline, ensuring proportional contributions from all disciplines before fitting the global ARIMA model. This improves the numerical stability of the model by incorporating information across disciplines. The forecast distributions are shown in @fig-grad-forecasts, with the mean forecast represented by the solid line and 90% prediction intervals indicated by the shaded area.

```{r}
#| label: fig-grad-forecasts
#| fig-cap: "Forecast of $G_{i,t}$: the number of graduates by discipline, 2024--2035, based on historical data from 2006--2023. The shaded regions represent the 90% prediction intervals, and the solid lines indicate the mean estimates."
tar_read(fig_grad_forecasts)
```

## Death probabilities

```{r}
#| label: fig-death-probs-forecasts
#| fig-cap: "Forecasts of $q_{x,t}$: age-specific probabilities of death (on a logarithmic scale) for 2035, based on historical data from 1971--2021. The shaded regions represent the 90% prediction intervals, and the solid lines indicate the mean estimate."
tar_read(fig_model_mxt3)
```

The death probabilities shown in @fig-death-probs where first smoothed using the partially monotonic penalized spline approach of @HU07. Then the functional data model @eq-fdm was estimated, with ARIMA models fitted to the coefficients. The forecasts for one year are shown in @fig-death-probs-forecasts, with the mean forecast represented by the solid line and 90% prediction intervals indicated by the shaded area. Note that the historical data (shown in gray) represent unsmoothed values, while the forecasts are based on the smoothed functional data model. The additional variation seen in the historical data is captured in the model through the Binomial death process.

## Remainder

The remainder, $E_{i,x,t}$, is also modelled using a functional data model [@HU07], with stationary ARMA models fitted to the principal component scores. The scores were found to be stationary using the KPSS test [@KPSS]. The forecasts for one year are shown in @fig-net-migration-disciplines, with the mean forecast represented by the solid line and 90% prediction intervals indicated by the shaded area.

```{r}
#| label: fig-net-migration-disciplines
#| fig-cap: "Forecasts of $E_{i,x,t}$: the remainder by discipline for 2035, based on historical data from 2006--2020. The shaded regions represent the 90% prediction intervals, and the solid lines indicate the mean estimates."
tar_read(fig_model_Ext3)
```

## Simulating future populations

We use the demographic growth-balance model @eq-dgb2 to iteratively simulate future populations, using the models described above for the components. The following steps outline the process.

A total of 1000 simulations are run to obtain a distribution of future age-specific population scenarios. The average of the 1000 simulations provides the mean age-specific forecast, while quantiles estimate forecast uncertainty. @fig-forecast-working-population presents the mean and 90% prediction intervals for 2035.

```{r}
#| label: fig-forecast-working-population
#| fig-cap: "Forecasts of $P_{i,x,t}$: the working population by discipline for 2035. The shaded regions represent the 90% prediction intervals, and the solid lines indicate the mean estimates."
#| fig-height: 5.5
# 2025 and 2035
tar_read(fig_Pxt_future_discipline) +
  ggplot2::labs(title = "Forecasts of working population by discipline: 2035")
```

In 2035, forecast variability is highest in the age period 20--35 years, before gradually narrowing as the workforce ages. This is due to the relatively high uncertainty in the new graduates component compared to the other components. Mid-to-late career estimates primarily reflect the aging of existing cohorts. The prediction intervals becomes especially narrow during the retirement phase, where the workforce dynamics become more predictable. Since retirements increase after the late 50s, workforce participation beyond 60 serves as a benchmark for identifying trends in delayed retirement and extended career duration. Over the next ten years, we expect an aging workforce in all but the Mathematical Sciences, where a large increase in the population is forecast.

```{r}
#| label: fig-forecast-total-working-scientists
#| fig-cap: "Forecasted total number of working scientists across scientific disciplines from 2022 to 2041. The shaded region represents the 90% prediction interval, the coloured line indicates the mean estimate, and the black line represents historical data."
tar_read(fig24)
```

Also visible in @fig-forecast-working-population are cohort effects, where fluctuations in earlier ages and years propagate through to later ages and years. This is particularly evident in the mid-career years because there are few deaths and retirements, few graduates older than 30, and the variation due to the remainder term is relatively small after age 30.

Summing over ages allows for estimating the forecast distribution of the total number of working scientists in each future year, as shown in @fig-forecast-total-working-scientists. The forecasts indicate continued growth, but at a gradually slower pace for all disciplines other than the Mathematical Sciences. Where the lower bound is nearly flat, workforce stagnation is possible in a conservative scenario. Even in the optimistic scenario (corresponding to the upper bound), growth only slightly exceeds the current pace, except for the Mathematical Sciences. As noted earlier, the divergent behaviour of the Mathematical Sciences is likely driven by the growing importance of data science and related fields.

The wide prediction intervals reflect the uncertainty in the forecasts, and show that caution is needed when interpreting the results. The only discipline where there is clear evidence of growth or decline is the Mathematical Sciences. For all other disciplines, the prediction intervals include the current level, indicating that stagnation, increase, or decline is possible.

# Discussion {#sec-conclusion}

While these forecasts provide a foundation for workforce planning, it is important to note that they are entirely driven by historical trends and do not account for possible new developments, such as the impact of AI and other emerging technologies on the labour market in different scientific disciplines. Other factors, such as policy changes or global economic shifts, may also influence workforce trends and should be taken into account.

Things that could be improved:

- Include backtesting results by fitting the model to earlier periods and comparing forecasts to actual outcomes
- Discuss forecast accuracy metrics and how they compare across disciplines
- Address the challenge of validating long-term demographic forecasts with limited historical data
- Examine how sensitive forecasts are to key assumptions (e.g., constant retirement rates across disciplines, independence of components)
- Discuss the impact of the choice to use general population mortality rates rather than discipline-specific rates
- Analyze how changes in the each component affect overall forecasts
- Address potential interventions that could alter the forecasted trajectories (e.g., targeted recruitment, retention programs)
- Discuss alternative methods.

- Provide more detailed discussion of the sources of uncertainty in your forecasts

- Discuss whether the methodology could be applied to other countries or workforce sectors
- Address unique aspects of the Australian context that might limit generalizability

- Expand beyond AI to discuss other technological disruptions (automation, quantum computing, biotechnology)
- Consider demographic shifts (changing education patterns, delayed career entry, career mobility)
- Discuss potential "black swan" events that could dramatically alter workforce dynamics

- Discuss the implications of using census data with 5-year gaps
- Address the challenge of tracking career changes and international mobility
- Suggest improvements to data collection that would enhance future forecasting efforts

# Software and reproducibility

All results presented here can be reproduced using the code available at <https://github.com/robjhyndman/age_structure_forecasts>. The analysis was conducted using R version 4.5.1 [@R], with the following R packages: vital [@Rvital], tsibble [@Rtsibble], fable [@Rfable], targets [@targets;@targetspaper], ggplot2 [@ggplot2;@ggplotbook], and other tidyverse [@tidyverse] packages.

# Acknowledgements

We thank Alexandra Lucchetti from the Australian Academy of Science for sourcing the data required for this project, and for helpful feedback on earlier versions of this report.