### 2025-11-23 asf1.tex
Added diff

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
doublespace,
  times]{anzsauth}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{setspace}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage[]{natbib}
\bibliographystyle{anzsj}

\usepackage[section]{placeins}
\usepackage{colortbl,todonotes}
\usepackage{orcidlink}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Forecasting the age structure of the scientific workforce in Australia},
  pdfauthor={Rob J Hyndman; Khuyen Vanh Nguyen},
  pdfkeywords={cohort analysis, demographic modelling, functional data
models, labour market, workforce planning},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\usepackage{moreverb}
\usepackage{url}
\usepackage{grffile}
\usepackage[UKenglish]{isodate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% The year in the following may need to be updated!
\def\volumeyear{}

% The following command ("obviously") effects line numbering
% of the document.

\usepackage{lineno}
\linenumbers

\def\firstletters{\bgroup \catcode`-=10 \catcode`(=10 \filA}
\def\filA#1{\filB#1 {\end} }
\def\filB#1#2 {\ifx\end#1\egroup \else#1 \expandafter\filB\fi}

\runningheads{
Forecasting the workforce age structure
}{
\firstletters{ROB J}HYNDMAN, AND  \firstletters{KHUYEN VANH}NGUYEN 
}

\title{Forecasting the age structure of the scientific workforce in
Australia}

\author{
Rob J Hyndman\addressnum{1} and 
Khuyen Vanh Nguyen
\addressnum{1}
}
\affiliation{
Monash University
}

\address{
\addressnum{1} Department of Econometrics \& Business Statistics, Monash
University\\
\hspace*{1ex} Email: \texttt{Rob.Hyndman@monash.edu} 
}

\date{}
\begin{document}

\begin{abstract}
Planning for a future workforce requires forecasts of age structure
changes to inform policy decisions, particularly related to universities
and immigration. We modify the traditional demographic growth-balance
equation to account for workforce-specific dynamics, replacing births
with graduate entry, modelling exits through death and retirement, and
including a remainder term that captures migration and career changes.
Functional data models are used to model age-specific components, while
ARIMA models are used for time series components. Simulation is employed
to generate forecast distributions, capturing uncertainty from all
components. The approach is illustrated using data on Australia's
scientific workforce, allowing us to forecast the age distribution of
various scientific disciplines for the next ten years. This analysis was
central to an Australian Academy of Science initiative examining the
capability of Australia's science system and identifying workforce gaps.
\end{abstract}

\keywords{cohort analysis; demographic modelling; functional data
models; labour market; workforce planning}

\ack{We thank Alexandra Lucchetti from the Australian Academy of Science
for sourcing the data required for this project, and for helpful
feedback on earlier versions of this paper. Rob Hyndman is a member of
the Australian Research Council Industrial Transformation Training
Centre in Optimisation Technologies, Integrated Methodologies, and
Applications (OPTIMA), Project ID IC200100009. He also receives funding
from the Australian Research Council through Discovery Project
DP250100702.}

\maketitle

\setstretch{1.5}
\section{Introduction}\label{sec-intro}

In planning for the future labour market, it is necessary to forecast
the age structure of the workforce in order to enable informed
decision-making on policies, especially concerning universities and
immigration. We propose a statistical modelling approach to this
problem, illustrated using various scientific disciplines in Australia,
forecasting future workforce age structures over the next decade. The
forecasts described have been used by the Australian Academy of Science
as part of \emph{Australian Science, Australia's Future: Science 2035},
an initiative assessing the capability of the national science system
and its role in achieving Australia's ambitions \citep{AAS2035}.

The economic implications of workforce age structure shifts are
well-documented \citep[e.g.,][]{Bloom2007}, affecting productivity,
pensions and superannuation, and skill shortages
\citep{PC2013, OECD2019a, OECD2019b, pension}. The social implications
are also significant, with an aging workforce leading to changes in
workplace dynamics, potential problems with intergenerational knowledge
transfer, and the need for policies that support older workers. Yet this
problem does not appear to have been previously addressed from a
statistical modelling perspective.

Our approach builds on functional data models, introduced to demographic
modelling by \citet{HU07}. They combined nonparametric smoothing and
functional principal components for age-specific demographic rates.
These models were then used by \citet{HB08} for mortality, fertility,
and migration rates, providing stochastic data generating processes for
the components of demographic balance equations. These separate
component models were then simulated to form future sample paths,
leading to age- and sex-specific stochastic population forecasts. The
modelling framework was later extended by \citet{HBY13} to ensure
coherence of forecasts between sexes or other demographic groups.

We propose to adapt this framework to workforce dynamics by redefining
the demographic components in two ways. First, we replace fertility with
workforce entry, which functions more like a migration process than a
birth process because graduates can enter the workforce at any age.
Second, we allow workers to leave the workforce through two processes:
retirement and death. Of course, people may also leave the workforce for
other reasons, such as a career change or family commitments, but since
we do not have data on these processes, we model them implicitly via a
remainder term.

We describe the methodology in Section~\ref{sec-methodology}. By way of
illustration, we apply the methodology to major scientific disciplines
in Australia, focusing on the Natural and Physical Sciences. We describe
the data sources in Section~\ref{sec-data}, with the results provided in
Section~\ref{sec-results}. The aim of this analysis is to inform future
workforce planning and policy decisions to support the growth of
Australia's scientific community. Finally, we provide some discussion
and conclusions in Section~\ref{sec-conclusion}.

\section{Methodology}\label{sec-methodology}

Suppose our workforce is divided into \(I\) groups, indexed by
\(i=1,\dots,I\). In our application, these are scientific disciplines,
but in principle they could refer to any subdivision of workers. Let
\(P_{i,x,t}\) denote the number of equivalent full-time workers in group
\(i\) who are aged \(x\) at the start of year \(t\), where
\(x=15,16,\dots\). The starting age of 15 is chosen because it is the
minimum age at which individuals are counted as part of the labour force
in the Australian Census \citep{ABSLFSP}. We assume that data are
available for years \(t=1,\dots,T\), and that forecasts are required for
\(P_{i,x,T+h}\) across all ages and groups, for some forecast horizon
\(h>0\).

People can leave the workforce of a group through death, retirement,
emigration, or career change; they can enter the workforce through
graduation, immigration, or career change. Unfortunately, we typically
do not have data on many of these processes, so we will combine career
change, emigration and immigration into a remainder term, which we
denote as \(E_{i,x,t}\). Let \(D_{i,x,t}\) denote the number of deaths
of workers in group \(i\) of age \(x\) in year \(t\), \(R_{i,x,t}\)
denote the number of retirements from the same group of workers, and
\(G_{i,x,t}\) denote the number of new graduates of age \(x\) in year
\(t\) who take up work in group \(i\). The numbers in each case are for
people aged \(x\) at the \emph{start} of year \(t\). Then population
changes can be described using a model similar to the stochastic
population model of \citet{HB08}, excluding the birth process:
\begin{equation}\phantomsection\label{eq-dgb}{
P_{i,x+1,t+1} = P_{i,x,t} - D_{i,x,t} - R_{i,x,t} + G_{i,x,t} + E_{i,x,t},
}\end{equation} where

\begin{itemize}
\tightlist
\item
  \(D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{i,x,t})\), with
  \(q_{i,x,t}\) being the probability of death for group \(i\) at age
  \(x\) in year \(t\); and
\item
  \(R_{i,x,t} \sim \text{Binomial}((P_{i,x,t}-D_{i,x,t}), r_{i,x,t})\),
  with \(r_{i,x,t}\) being the probability of retirement from group
  \(i\) at age \(x\) in year \(t\).
\end{itemize}

That is, the population each year is equal to the population from the
previous year having aged 1 year, minus the deaths or retirements that
occurred during the previous year, plus the new graduates, plus any
other changes due to migration or career change (which may be negative).
We assume that \(E_{i,x,t} = G_{i,x,t} = 0\) above some age threshold
(say \(x=100\)). Once \(P_{i,x,t} = 0\) when \(x\) is above that
threshold, all future populations \(P_{i,x+k,t+k} = 0\), for
\(k=1,2,\dots\). That is, when the cohort aged \(x\) in year \(t\) has
all retired or died, and \(x\) is above the threshold, they will not be
replaced by new workers of the same age.

As a first approximation, the components \(q\), \(r\), \(E\) and \(G\)
can be assumed to behave independently for each combination of \(i\),
\(x\) and \(t\). In reality, there may be some negative correlation
between \(G\) and \(E\) as insufficient graduates would probably lead to
employers finding people from overseas, while too many graduates would
lead to scientists seeking work elsewhere.

The choice of a Binomial rather than a Poisson distribution (in contrast
to \citet{Brillinger1986}) for deaths and retirements is because the
Binomial distribution ensures that the number of deaths and retirements
cannot exceed the population at risk. In a simulation context, with very
small populations, this is important to avoid nonsensical results.

It is unlikely that we have available separate death and retirement
counts for each group, and retirement data is not available in all
years. So we will let \(q_{i,x,t} = q_{x,t}\) and \(r_{i,x,t} = r_{x}\),
assuming that death rates and retirement rates are the same across all
groups, and that retirement rates do not change over time. Similarly,
graduation numbers are rarely available by discipline and age, so we
will approximate \(G_{i,x,t} = g_xG_{i,t}\) where \(G_{i,t}\) is the
total number of graduates in year \(t\) and \(g_x\) is the proportion of
graduates by age across all disciplines.

This leads to the simpler model
\begin{equation}\phantomsection\label{eq-dgb2}{
P_{i,x+1,t+1} = P_{i,x,t} - D_{i,x,t} - R_{i,x,t} + g_xG_{i,t} + E_{i,x,t},
}\end{equation} where

\begin{itemize}
\tightlist
\item
  \(D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})\); and
\item
  \(R_{i,x,t} \sim \text{Binomial}((P_{i,x,t}-D_{i,x,t}), r_{x})\).
\end{itemize}

For the age-specific time-varying components \(q_{x,t}\) and
\(E_{i,x,t}\), we will use functional data models \citep{HU07}. For
\(q_{x,t}\), this is of the form
\begin{equation}\phantomsection\label{eq-fdm}{
\log(q_{x,t}) = \mu(x) + \sum_{k=1}^K \beta_{k,t} \phi_k(x) + \varepsilon_{x,t},
}\end{equation} where \(\mu(x)\) is the mean function, \(\phi_k(x)\) are
functional principal components, \(\beta_{k,t}\) are the principal
component scores, and \(\varepsilon_{x,t}\) is an error term. Each
\(\beta_{k,t}\) is then modelled using a univariate time series model,
such as an ARIMA model. A log-link function is used to ensure that the
probabilities remain positive. An inverse logit link function could also
be used, if the probabilities are close to 1 for some ages. A similar
model is used for \(E_{i,x,t}\), with separate mean functions and
principal components for each group \(i\). The number of components
\(K=6\), for the reasons outlined in \citet{HB08}.

Functional data models have been widely used in demography and other
fields, and have been shown to work particularly well for age-specific
demographic processes \citep{HB08, LCcomparison}. They enable the
inherent smoothness over age to be captured, while modelling the
autocorrelation over time using relatively simple univariate time series
models applied to the principal component scores. In our application, we
use univariate ARIMA models for the \(q_{x,t}\) scores, and ARMA models
for the \(E_{i,x,t}\) scores. The assumption of stationarity for the
\(E_{i,x,t}\) scores is validated for the disciplines we consider.

For the time-varying component, \(G_{i,t}\), we use a global ARIMA model
\citep{HM21} to capture the dynamics over time and across disciplines.
The global model pools information across disciplines to improve
forecast accuracy, especially for disciplines with limited historical
data.

To forecast future working population numbers, \(P_{i,x,t}\), \(t>T\),
we simulate future sample paths of each of the components \(G_{i,t}\),
\(q_{x,t}\), and \(E_{i,x,t}\), simulate \(D_{i,x,t}\) and \(R_{i,x,t}\)
from their respective Binomial distributions, and then use the
demographic growth-balance equation Equation~\ref{eq-dgb2} iteratively
to obtain \(P_{i,x,t}\) for \(t=T+1,T+2,\dots\). This simulation-based
approach allows us to capture the uncertainty in each of the components,
leading to a distribution of possible future outcomes for \(P_{i,x,t}\).

This model is somewhat pragmatic given the data available in our
specific application. If better data were available, other variations on
Equation~\ref{eq-dgb} could be used. For example, if death rates were
available by discipline, then we would replace \(q_{x,t}\) by
\(q_{i,x,t}\) in the Binomial deaths distribution. If retirement rates
were available by year, or by discipline, we could similarly replace
\(r_x\) by a more specific retirement rate in the Binomial retirements
distribution. If we had data on graduations by age and discipline, we
could replace \(g_xG_{i,t}\) by \(G_{i,x,t}\). If we had data on
migration, we could split the remainder \(E_{i,x,t}\) into several
components, and model them separately. None of this changes the overall
modelling framework we are proposing.

\section{Data}\label{sec-data}

To illustrate the methodology, we consider the Natural and Physical
Sciences as defined in the Australian Standard Classification of
Education (ASCED) by the \citet{ABSASCED}. We refer to ASCED's Narrow
Fields as ``disciplines''; these comprise Physics and Astronomy,
Mathematical Sciences, Chemical Sciences, Earth Sciences, Biological
Sciences, Other Natural and Physical Sciences, and Natural and Physical
Sciences not further defined (n.f.d.). Table~\ref{tbl-disciplines} lists
the detailed fields within each scientific discipline.

\begin{table}

\caption{\label{tbl-disciplines}Classification of scientific
disciplines, based on the ASCED Narrow Fields of Education within the
Broad Field of Natural and Physical Sciences. The table lists their
corresponding Detailed Fields. ``n.e.c.'' stands for ``Not Elsewhere
Classified.''}

\centering{

\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{9cm}}
\toprule
\textbf{Narrow Fields} & \textbf{Detailed Fields}\\
\midrule
Physics and Astronomy & Physics, Astronomy.\\
Mathematical Sciences & Mathematics, Statistics, Mathematical Sciences, n.e.c.\\
Chemical Sciences & Organic Chemistry, Inorganic Chemistry, Chemical Sciences, n.e.c.\\
Earth Sciences & Atmospheric Sciences, Geology, Geophysics, Geochemistry, Soil Science, Hydrology, Oceanography, Earth Sciences, n.e.c.\\
Biological Sciences & Biochemistry and Cell Biology, Botany, Ecology and Evolution, Marine Science, Genetics, Microbiology, Human Biology, Zoology, Biological Sciences, n.e.c.\\
Other Natural and Physical Sciences & Medical Science, Forensic Science, Food Science and Biotechnology, Pharmacology, Laboratory Technology, Natural and Physical Sciences, n.e.c.\\
\bottomrule
\end{tabular}}

}

\end{table}%

We define the population of workers in a discipline as those who are
active in the labour market and hold a bachelor's degree or higher in
that discipline. For the purposes of this analysis, we will omit ``Other
Natural and Physical Sciences'' and ``Natural and Physical Sciences
n.f.d.''.

\subsection{Working population}\label{working-population}

Data on the working population were sourced from the \emph{Census of
Population and Housing} \citep{ABSCensus} for census years 2006, 2011,
2016, and 2021. This dataset encompasses one-year age groups, the
highest level of completed non-school qualification level (QALLP), the
corresponding field of study \citep[QALFP,][]{ABSQALFP}, and the
industries in which individuals work. However, labour force
participation status \citep{ABSLFFP} is available only for 2016 and
2021. To estimate worker numbers for 2006 and 2011, the average
participation rates from 2016 and 2021 were applied, assuming overall
age distributions remain consistent.

The resulting estimates of the number of scientists who are active in
the Australian labour market is shown in
Figure~\ref{fig-cohort-interpolation} as the thick lines. Cohort
interpolation \citep{STUPP88}, applying linear interpolation within each
age cohort between census years, is used to estimate values for the
intercensal years (shown as thin lines), giving \(P_{i,x,t}\) for each
discipline \(i\), age \(x\), and year \(t\).

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-cohort-interpolation-1.pdf}}

}

\caption{\label{fig-cohort-interpolation}\(P_{i,x,t}\): Estimated number
of working scientists in Australia by discipline and age, 2006--2021.
Thicker lines are used to denote census years.}

\end{figure}%

\subsection{Retirements}\label{retirements}

Retirement data was sourced from the \emph{Retirement and Retirement
Intentions} dataset (Catalogue 6238) for the 2022--2023 financial year
\citep{ABS6238}. The data are categorised by the industry of an
individual's main job, and are provided in four broad age groups
(45--59, 60--64, 65--69 and 70+). There are 19 industry categories, with
the largest numbers of scientists working in Education and Training
(15.8\%), Professional, Scientific and Technical Services (15.5\%), and
Health Care and Social Assistance (14.6\%). The proportions in other
industries are much smaller. We take a weighted average of retirement
intentions using these top three industries, with proportions rescaled
to sum to 1. The resulting values are shown in
Figure~\ref{fig-retirements} as the gray line. To obtain a
single-year-of-age retirement distribution, we disaggregate the data
using a monotonic cubic spline applied to the cumulative values of these
age groups \citep{SHW04}. The resulting smoothed distribution (\(r_x\))
is shown as the black line in Figure~\ref{fig-retirements}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-retirements-1.pdf}}

}

\caption{\label{fig-retirements}\(r_x\): Age distribution of retirement
intentions, based on data from the 2022--2023 Australian financial year.
The grey line shows the age-group probabilities; the black line shows
the smoothed probabilities.}

\end{figure}%

\subsection{Deaths}\label{deaths}

Age-specific mortality rates from 1971 to 2021 were obtained from the
\citet{HMD}. Using standard life table methods, these rates are
converted into age-specific probabilities of death, as shown in
Figure~\ref{fig-death-probs}. Over time, mortality probabilities have
generally declined across all age groups, reflecting improvements in
Australian life expectancy.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-death-probs-1.pdf}}

}

\caption{\label{fig-death-probs}\(q_{x,t}\): Age-specific probabilities
of death (on a logarithmic scale) for each year from 1971 to 2021.}

\end{figure}%

No data are available for specific industry groups, so we assume that
all scientists have the same mortality probabilities as the general
population. These probabilities serve as estimates of \(q_{x,t}\).

\subsection{Graduate completions}\label{graduate-completions}

Graduate completion statistics were obtained from the \emph{Award Course
Completions} dataset \citep{DeptEdu}. Figure~\ref{fig-completions} shows
the distribution of graduate completions with a bachelor's degree or
higher, by age for each year from 2006 to 2023. Some missing values
result in gaps in certain lines, but the overall pattern remains highly
consistent across years. Given this consistency, the data is averaged
across all available years, and then smoothed by applying monotonic
cubic splines to the cumulative values \citep{SHW04}. The resulting
averaged distribution, shown as the black line in
Figure~\ref{fig-completions}, smooths out year-to-year fluctuations and
provides an estimate of \(g_x\).

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-completions-1.pdf}}

}

\caption{\label{fig-completions}\(g_x\): Estimated distribution of
graduate completions by age (black). This is estimated by averaging and
smoothing the data for the years 2006 to 2023 (coloured).}

\end{figure}%

The Department of Education provides data on the number of graduates
with a bachelor's degree or higher, categorised by discipline and year
\citep{DeptEdu_Private}. This dataset includes both domestic and
international students. The total number of graduates, \(G_{i,t}\), in
each discipline \(i\) and year \(t\), are shown in
Figure~\ref{fig-graduates}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-graduates-1.pdf}}

}

\caption{\label{fig-graduates}\(G_{i,t}\): Total number of graduates
with a bachelor's degree or higher by discipline from 2006 to 2023.}

\end{figure}%

The large increase in the working population observed in the 2021 Census
for Mathematical Sciences (Figure~\ref{fig-cohort-interpolation}) can be
partly attributed to the sharp rise in graduate numbers between 2016 and
2021. This is probably due to the impact of data science, and the
growing importance of statistics and machine learning in many areas of
employment.

\subsection{Remainder}\label{remainder}

The demographic growth-balance equation (Equation~\ref{eq-dgb2}), when
rearranged, provides an expression for the remainder including net
migration and career changes:
\begin{equation}\phantomsection\label{eq-remainder}{
E_{i,x,t} = P_{i,x+1,t+1} - P_{i,x,t} - D_{i,x,t} - R_{i,x,t} - g_xG_{i,t},
}\end{equation} However, we do not have data on \(D_{i,x,t}\) and
\(R_{i,x,t}\), so we replace these by their expected values,
\(P_{i,x,t} q_{x,t}\) and \(P_{i,x,t}(1-q_{x,t}) r_x\), respectively. We
can only estimate remainders up to 2020 because we need data for both
year \(t\) and year \(t+1\) in Equation~\ref{eq-remainder}, and our
working population data only extends to 2021. The estimated remainders
are shown in Figure~\ref{fig-remainder}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-remainder-1.pdf}}

}

\caption{\label{fig-remainder}Estimated remainder \(E_{i,x,t}\) by
discipline, age and year (2006--2020).}

\end{figure}%

The inclusion of international students in the graduate data leads to
large positive values of the remainder for the teenage years, followed
by large negative values when these students return to their home
countries after graduation.

\section{Results}\label{sec-results}

\subsection{Graduate completions}\label{graduate-completions-1}

To forecast future graduate numbers, \(G_{i,t}\), a global ARIMA model
was employed, following the principles outlined by \citet{HM21}. The
global model captures overall trends across disciplines by scaling
graduate data within each discipline, ensuring proportional
contributions from all disciplines before fitting the global ARIMA
model. This improves the numerical stability of the model by
incorporating information across disciplines. The forecast distributions
are shown in Figure~\ref{fig-grad-forecasts}, with the mean forecast
represented by the solid line and 90\% prediction intervals indicated by
the shaded area.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-grad-forecasts-1.pdf}}

}

\caption{\label{fig-grad-forecasts}Forecast of \(G_{i,t}\): the number
of graduates by discipline, 2024--2035, based on historical data from
2006--2023. The shaded regions represent the 90\% prediction intervals,
and the solid lines indicate the mean estimates.}

\end{figure}%

\subsection{Death probabilities}\label{death-probabilities}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-death-probs-forecasts-1.pdf}}

}

\caption{\label{fig-death-probs-forecasts}Forecasts of \(q_{x,t}\):
age-specific probabilities of death (on a logarithmic scale) for 2035,
based on historical data from 1971--2021. The shaded regions represent
the 90\% prediction intervals, and the solid lines indicate the mean
estimate.}

\end{figure}%

The death probabilities shown in Figure~\ref{fig-death-probs} were first
smoothed using the partially monotonic penalised spline approach of
\citet{HU07}. Then the functional data model Equation~\ref{eq-fdm} was
estimated, with ARIMA models fitted to the coefficients. The forecasts
for one year are shown in Figure~\ref{fig-death-probs-forecasts}, with
the mean forecast represented by the solid line and 90\% prediction
intervals indicated by the shaded area. Note that the historical data
(shown in gray) represent unsmoothed values, while the forecasts are
based on the smoothed functional data model. The additional variation
seen in the historical data is captured in the model through the
Binomial death process.

\subsection{Remainder}\label{remainder-1}

The remainder, \(E_{i,x,t}\), is also modelled using a functional data
model \citep{HU07}, with stationary ARMA models fitted to the principal
component scores. The scores were found to be stationary using the KPSS
test \citep{KPSS}. The forecasts for one year are shown in
Figure~\ref{fig-net-migration-disciplines}, with the mean forecast
represented by the solid line and 90\% prediction intervals indicated by
the shaded area.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-net-migration-disciplines-1.pdf}}

}

\caption{\label{fig-net-migration-disciplines}Forecasts of
\(E_{i,x,t}\): the remainder by discipline for 2035, based on historical
data from 2006--2020. The shaded regions represent the 90\% prediction
intervals, and the solid lines indicate the mean estimates.}

\end{figure}%

\subsection{Simulating future
populations}\label{simulating-future-populations}

We use the demographic growth-balance model Equation~\ref{eq-dgb2} to
iteratively simulate future populations, using the models described
above for the components. The following steps outline the process.

A total of 1000 simulations are run to obtain a distribution of future
age-specific population scenarios. The average of the 1000 simulations
provides the mean age-specific forecast, while quantiles estimate
forecast uncertainty. Figure~\ref{fig-forecast-working-population}
presents the mean and 90\% prediction intervals for 2035.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-forecast-working-population-1.pdf}}

}

\caption{\label{fig-forecast-working-population}Forecasts of
\(P_{i,x,t}\): the working population by discipline for 2035. The shaded
regions represent the 90\% prediction intervals, and the solid lines
indicate the mean estimates.}

\end{figure}%

In 2035, forecast variability is highest in the age period 20--35 years,
before gradually narrowing as the workforce ages. This is due to the
relatively high uncertainty in the new graduates component compared to
the other components. Mid-to-late career estimates primarily reflect the
aging of existing cohorts. The prediction intervals become especially
narrow during the retirement phase, where the workforce dynamics become
more predictable. Since retirements increase after the late 50s,
workforce participation beyond 60 serves as a benchmark for identifying
trends in delayed retirement and extended career duration. Over the next
ten years, we expect an aging workforce in all but the Mathematical
Sciences, where a large increase in the population is forecast.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{figures/fig-forecast-total-working-scientists-1.pdf}}

}

\caption{\label{fig-forecast-total-working-scientists}Forecasted total
number of working scientists across scientific disciplines from 2022 to
2041. The shaded region represents the 90\% prediction interval, the
coloured line indicates the mean estimate, and the black line represents
historical data.}

\end{figure}%

Cohort effects are also visible in
Figure~\ref{fig-forecast-working-population}, where fluctuations in
earlier ages and years propagate through to later ages and years. This
is particularly evident in the mid-career years because there are few
deaths and retirements, few graduates older than 30, and the variation
due to the remainder term is relatively small after age 30.

Summing over ages allows for estimating the forecast distribution of the
total number of working scientists in each future year, as shown in
Figure~\ref{fig-forecast-total-working-scientists}. The forecasts
indicate continued growth, but at a gradually slower pace for all
disciplines other than the Mathematical Sciences. Where the lower bound
is nearly flat, workforce stagnation is possible in a conservative
scenario. Even in the optimistic scenario (corresponding to the upper
bound), growth only slightly exceeds the current pace, except for the
Mathematical Sciences. As noted earlier, the divergent behaviour of the
Mathematical Sciences is likely driven by the growing importance of data
science and related fields.

The wide prediction intervals reflect the uncertainty in the forecasts,
and show that caution is needed when interpreting the results. The only
discipline where there is clear evidence of growth or decline is the
Mathematical Sciences. For all other disciplines, the prediction
intervals include the current level, indicating that stagnation,
increase, or decline is possible.

\section{Discussion}\label{sec-conclusion}

While these forecasts provide a foundation for workforce planning, it is
important to note that they are entirely driven by historical trends and
do not account for possible new developments, such as the impact of AI
and other emerging technologies on the labour market in different
scientific disciplines. Other factors, such as policy changes or global
economic shifts, may also influence workforce trends. These exogenous
factors could be accounted for by adding covariates into the time series
models used, provided relevant data are available.

Evaluating and validating these forecasts is challenging due to the
relatively long forecast horizon compared to the available historical
data. While time series cross-validation \citep{fpp3} could be used to
assess forecast accuracy for shorter horizons, the benefit of the
forecasts is primarily for longer horizons, where such validation is not
possible. The uncertainty in the forecasts, as reflected in the wide
prediction intervals, highlights the need for caution when interpreting
the results.

If more detailed data were available, the model could be refined further
by including, for example, discipline-specific death rates, retirement
data by year and/or discipline, graduate data by age and discipline, and
data on migration and career changes. It is not clear how much these
refinements would improve forecast accuracy, but they would likely
reduce uncertainty in the forecasts.

Forecasts are often designed not just to predict the future, but also to
inform policy decisions, and so modify the future. In this context,
these forecasts could be used to identify potential skill shortages or
surpluses in specific disciplines, guiding decisions on university and
immigration policy, and thus changing the future outcomes
\citep{causality2022}. Consequently, forecast accuracy may be less
important than understanding the range of possible outcomes and their
implications for policy.

While this analysis has focused on the scientific workforce in
Australia, the methodology could be applied to other countries or
workforce sectors, provided similar data are available. The specific
components of the demographic growth-balance equation may need to be
adapted to reflect the available data in other applications.

\section{Software and
reproducibility}\label{software-and-reproducibility}

All results presented here can be reproduced using the code available at
\url{https://github.com/robjhyndman/age_structure_forecasts}. The
analysis was conducted using R version 4.5.1 \citep{R}, with the
following R packages: vital \citep{Rvital}, tsibble \citep{Rtsibble},
fable \citep{Rfable}, targets \citep{targets, targetspaper}, ggplot2
\citep{ggplot2, ggplotbook}, and other tidyverse \citep{tidyverse}
packages.

\bibliography{refs.bib}

\end{document}